Long-Term Reproducibility through Replication:\\ the ReScience Initiative

### 2017-06-08 draft.tex
Revision of the "reproduction and replication" section

With the increasing recognition of the replication crisis as a problem
for scientific research, it has become a major subject of debate, but
unfortunately no common terminology has emerged so far. One reason for
this diverse use of terms is that each field of research has its own
specific technical and social obstacles on the road to published
results and findings that can be verified by other scientists. Here we
briefly summarize the obstacles that arise from the use of computers
and software in scientific research, and introduce the terminology we
will use in the rest of this article.\\

\textit{Reproducing} the result of a computation means running the
same software on the same input data and obtaining the same results.
Obstacles to reproducibility are the loss of parts of the software or
input data, lack of a computing environment that is sufficiently
similar to the one used initially, and insufficient instructions for
making the software work. An obstacle specific to numerical
computations is the use of floating-point arithmetic, whose rules are
subject to slightly different interpretations by different compilers
and runtime support systems.

@@

Strictly speaking, reproducibility is defined in the context of
identical computational environments. However, useful scientific
software is expected to be robust with respect to certain changes in
this environment. A computer program that produces different results
when compiled using different compilers, or run on two different
computers, would be considered suspect by most practitioners, even if
it were demonstrably correct in one specific environment. Ultimately
it is not the software that is of interest for science, but the models
and methods that it implements. The software is merely a vehicle to
perform computations based on these models and methods. If results
depend on hard to control implementation details of the software,
their relation to the underlying models and methods becomes unclear.\\

\textit{Replicating} a published result means writing and then running
new software based on the description of a computational model or
method provided in the original publication, and obtaining results
which support the same scientific conclusions as the original
ones. The main obstacle to replicability is an incomplete or imprecise
description of the models and methods.

@@

Replication can be attempted both for reproducible and for
non-reproducible results. However, when an attempt to replicate
non-reproducible work fails, yielding results so different that the
scientific conclusions drawn from them are different, it can be very
difficult to identify the cause of the disagreement. Reproducibility
guarantees the existence of a precise and complete description of the
models and methods being applied in the original work, in the form of
software source code. The holy grail of computational science is
therefore a reproducible replication of reproducible original work.

>>>>

With the increasing recognition of the replication crisis as a problem
for scientific research, it has become a major subject of debate, but
unfortunately no common terminology has emerged so far. One reason for
the diverse use of terms is that each field of research has its own
specific technical and social obstacles on the road to published
results and findings that can be verified by other scientists. Here we
briefly summarize the obstacles that arise from the use of computers
and software in scientific research, and introduce the terminology we
will use in the rest of this article.\\

\textit{Reproducing} the result of a computation means running the
same software on the same input data and obtaining the same results.
The goal of a reproduction attempt is to verify that the computational
protocol leading to the results has been recorded correctly.
Performing computations reproducibly can be seen as a form of
provenance tracking, the software being a detailed record of all data
processing steps.

In theory, computation is a deterministic process and exact
reproduction should therefore be trivial. In reality, it is very
difficult to achieve because of the complexity of today's software
stacks and the tediousness of recording all interactions between a
scientist and a computer.  \citeauthor{Mesnard:2016} explain in
\citep{Mesnard:2016} how difficult it can be to reproduce a two-year
old computation even though all possible precautions were taken at the
time to ensure reproducibility.  The most frequent obstacles are the
loss of parts of the software or input data, lack of a computing
environment that is sufficiently similar to the one used initially,
and insufficient instructions for making the software work. An
obstacle specific to numerical computations is the use of
floating-point arithmetic, whose rules are subject to slightly
different interpretations by different compilers and runtime support
systems. A large variety of research practices and support tools have
been developed recently to facilitate reproducible computations. For a
collection of recipes that have proven useful, see
\citep{kitzes:2017}.

@@

Strictly speaking, reproducibility is defined in the context of
identical computational environments. However, useful scientific
software is expected to be robust with respect to certain changes in
this environment. A computer program that produces different results
when compiled using different compilers, or run on two different
computers, would be considered suspect by most practitioners, even if
it were demonstrably correct in one specific environment. Ultimately
it is not the software that is of interest for science, but the models
and methods that it implements. The software is merely a vehicle to
perform computations based on these models and methods. If results
depend on hard to control implementation details of the software,
their relation to the underlying models and methods becomes unclear
and unreliable.\\

\textit{Replicating} a published result means writing and then running
new software based on the description of a computational model or
method provided in the original publication, and obtaining results
that are similar enough to be considered equivalent. What exactly
``similar enough'' means strongly depends on the kind of computation
being performed, and can only be judged by an expert in the field.
The main obstacle to replicability is an incomplete or imprecise
description of the models and methods.

@@

Replication can be attempted both for reproducible and for
non-reproducible results. However, when an attempt to replicate
non-reproducible work fails, yielding results too different to be
considered equivalent, it can be very difficult to identify the cause
of the disagreement. Reproducibility guarantees the existence of a
precise and complete description of the models and methods being
applied in the original work, in the form of software source code,
which can be analyzed during the investigation of any
discrepancies. The holy grail of computational science is therefore a
reproducible replication of reproducible original work.
